# Set environment and import data
```{r}
rm(list=ls())
install.packages("rjags")
install.packages("ggplot2")
```

```{r}
wine=readRDS(file="./data/wine.Rda")
summary(wine)
```
# Something

```{r}
library("ggplot2")

```
We may think using both Pearson and Spearman rank correlation coefficient:
\begin{itemize}
  \item Pearson captures linear relationships
  \item Spearman captures monotonic relationships
\end{itemize}
```{r}
library(reshape2)
melted_wine <- melt(cor(wine))
head(melted_wine)
ggplot(data = melted_wine, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile(color = "white") + 
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") + theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
 coord_fixed()

```

```{r}
correlation_with_quality = tail(melted_wine, n=12)
correlation_with_quality["value"] = abs(correlation_with_quality["value"])
correlation_with_quality = correlation_with_quality[order(correlation_with_quality$value, decreasing = TRUE),]
correlation_with_quality
```

```{r}
correlation_with_quality$value

```

```{r}
boxplot(wine, las=3,main="Covariates boxplot",cex.axis=0.75)
```

```{r}
boxplot(wine$total.sulfur.dioxide, las=3,main="total sulfur dioxide boxplot",cex.axis=0.75)
```
```{r}
boxplot(wine$free.sulfur.dioxide, las=2,main="free sulfur dioxide boxplot", cex.axis=0.75)
```
```{r}
info = boxplot(wine$free.sulfur.dioxide, las=2,main="free sulfur dioxide boxplot", cex.axis=0.75, plot=FALSE)
outliers = info$out
outliers
```

```{r}
hist(wine$free.sulfur.dioxide, freq=TRUE, border="#55D8C1", col="#FCF69C")
```


```{r}
hist(wine$quality, freq=TRUE, border="#55D8C1", col="#FCF69C", plot=FALSE)
```

```{r}
length(wine$quality)
```
```{r}
table(unlist(wine$quality))
```


1.50 min for softmax video
2.02 I don't understand
Bayesian Learning Lecture 9 - to see how to write the predictive distribution-
ascii ~ = 126














